---
title: "Instructor Solution Key & Assessment Guide"
format:
  html:
    toc: true
    code-fold: show
---

This page contains the answers to all quizzes, homework assignments, and advanced expansion challenges for the 10-hour Text Analysis course.

---

## Module 1: Foundations & Tokenization

### Part A: Concept Quiz

**Q1. True or False:** The "Tidy Text" format is structured as one-document-per-row.
> **Answer:** **False.** The Tidy Text principle is **one-token-per-row**. This structure allows us to use standard data manipulation tools like `dplyr` to filter and join datasets at the word level.

**Q2. Multiple Choice:** Which specific R function is used to split a text column into individual tokens?
* A) `separate_rows()`
* B) `unnest_tokens()`
* C) `str_split()`
> **Answer:** **B) `unnest_tokens()`**. While `str_split` can do this, `unnest_tokens` is the dedicated tidytext function that handles lowercasing and punctuation removal automatically.

**Q3. True/False:** Tokenization always splits text strictly by white space.
> **Answer:** **False.** Good tokenizers (like those in `tidytext` or `nltk`) handle punctuation intelligently. For example, "It's" might be split into "It" and "'s" depending on the engine, and "New York" might be kept as one token if n-grams are used.

### Part B: Practical Homework (The Time Machine)

**Task:** Download H.G. Wells' *The Time Machine* and perform tokenization.

**R Solution:**
```{r}
#| eval: false
library(gutenbergr)
library(tidytext)
library(dplyr)

# ID 35 is The Time Machine
hg_wells <- gutenberg_download(35) %>% 
  unnest_tokens(word, text)

# Inspect the tidy structure
head(hg_wells)
```

**Python Solution:**

```{python}
#| eval: false
import nltk
# nltk.download('punkt') # Run once if needed

raw_text = "The Time Traveller (for so it will be convenient to speak of him)..."

# 1. Tokenize
tokens = nltk.word_tokenize(raw_text)

# 2. Filter for alphabetic words only (removing punctuation)
clean_words = [t for t in tokens if t.isalpha()]

print(clean_words[:10])
```


### Part C: Advanced Expansion Challenge

R Challenge: Count words appearing more than 50 times in The Time Machine.

```{r}
#| eval: false
hg_wells %>%
  count(word, sort = TRUE) %>%
  filter(n > 50)
```

Python Challenge: Tokenize by Sentence instead of word.

```{python}
#| eval: false
text_block = "The Time Traveller sat back. He smiled. It was done."
sentences = nltk.sent_tokenize(text_block)
print(sentences) 
# Output: ['The Time Traveller sat back.', 'He smiled.', 'It was done.']
```

## Module 2: Preprocessing

### Part A: Concept Quiz

Q1. Concept Check: What is the primary difference between Stemming and Lemmatization?

Answer: > * Stemming is a heuristic process that chops off the ends of words (e.g., "ponies" -> "poni"). It is fast but crude.

Lemmatization uses a dictionary and morphological analysis to find the actual root form (e.g., "ponies" -> "pony").

Q2. Concept Check: In the TF-IDF formula, what is the purpose of the IDF (Inverse Document Frequency) component?

Answer: To penalize words that appear in every document. If a word like "the" appears in all documents, its IDF score is near zero, lowering its overall weight.

Q3. R Knowledge: Which function converts a tidy data frame into a sparse matrix?

A) cast_sparse()

B) as.matrix()

Answer: A) cast_sparse(). This is crucial for connecting tidytext workflows to machine learning packages like topicmodels or glmnet.

### Part B: Advanced Practice

R Challenge: Visualize the top 10 words in Jane Austen's Persuasion.

```{r}
#| eval: false
library(ggplot2)

tidy_books %>%
  filter(book == "Persuasion") %>%
  count(word, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = reorder(word, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top Words in Persuasion")
```

Python Challenge: Demonstrate Stemming using the Porter Stemmer.

```{python}
#| eval: false
from nltk.stem import PorterStemmer
stemmer = PorterStemmer()

words = ["running", "runs", "ran", "runner"]
stemmed = [stemmer.stem(w) for w in words]

print(stemmed)
# Notice that "ran" typically stays "ran" in stemming, 
# whereas a lemmatizer might map it to "run".
```

## Module 3: Topic Modeling

### Part A: Concept Quiz

Q1. True/False: LDA is a supervised learning algorithm.

Answer: False. LDA is unsupervised. We do not tell it what the topics are in advance; the algorithm discovers the clusters based on word co-occurrence.

Q2. Parameter Check: What controls the number of topics the model finds?

A) Alpha

B) Beta

C) k

Answer: C) k. You must specify k (e.g., k=2 or k=5) when initializing the model.

### Part B: Practical Homework (Interpreting Topics)

Task: Run LDA with k=3 and interpret the results.

R Solution:

```{r}
#| eval: false
# Fit model with 3 topics
lda_3 <- LDA(dtm, k = 3, control = list(seed = 1234))

# Extract words
terms(lda_3, 5)
```
Interpretation Example:

Topic 1: "love, heart, dear" -> Label: Romance

Topic 2: "money, pounds, estate" -> Label: Wealth

Topic 3: "time, house, day" -> Label: Daily Life

### Part C: Advanced Expansion Challenge

R Challenge: Inspect the Gamma (γ) matrix to see document probabilities.

```{r}
#| eval: false
# The 'gamma' matrix shows the probability that a specific document belongs to a topic
gamma_results <- tidy(lda_model, matrix = "gamma")

# Sort to find which document is most strongly associated with Topic 1
gamma_results %>% 
  filter(topic == 1) %>%
  arrange(desc(gamma))
```


Python Challenge: Train a Gensim model with 3 topics.

```{python}
#| eval: false
lda_model_3 = models.LdaModel(
    bow_corpus, 
    num_topics=3, # Changed from 2 to 3
    id2word=dictionary, 
    passes=10
)

print(lda_model_3.print_topics())
```

## Module 4: Applications & Inference


### Part A: Concept Quiz

Q1. Score Range: The VADER compound sentiment score ranges from:

A) 0 to 100

B) -1 to +1

Answer: B) -1 to +1, where -1 is most negative, 0 is neutral, and +1 is most positive.

Q2. Methodology: TextBlob and VADER are examples of what type of sentiment analysis?

A) Deep Learning

B) Lexicon-based (Rule-based)

Answer: B) Lexicon-based. They rely on pre-built dictionaries of words rated for sentiment, rather than training a neural network from scratch.

### Part B: Practical Challenge (The "Bad" Review)

Task: Analyze a negative sentence to see if VADER detects it.

Python Solution:

```{python}
#| eval: false
from nltk.sentiment.vader import SentimentIntensityAnalyzer

sid = SentimentIntensityAnalyzer()
text = "I absolutely hated this product. It was a complete waste of time and money."
score = sid.polarity_scores(text)

print(f"Compound Score: {score['compound']}")
# Expected Result: A score close to -0.8 (Highly Negative)
```


### Part C: Advanced Expansion Challenge

R Challenge: Use the "Bing" lexicon for binary sentiment counts.

```{r}
#| eval: false
# Assumes hg_wells is loaded
bing_counts <- hg_wells %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment)

print(bing_counts)

```

Python Challenge: Compare VADER with TextBlob polarity.

```{python}
#| eval: false
from textblob import TextBlob

text = "The movie was not bad at all."
blob = TextBlob(text)

print(f"TextBlob Polarity: {blob.sentiment.polarity}")
# TextBlob handles negation ("not bad") to produce a positive score (> 0).
```